---
title: 'Base R: Smoothing Methods'
author: "David J Jackson"
date: "8/10/2020"
output: 
  html_document: 
    fig_width: 9
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### In the interest of Full disclosure ...
#### There are packages that have moivng average function build in:

* Zoo, forecast,TTR and pracma.
* There are three ways of adding rolling means to you code:
* Embed the code directly
* Use the source command to read a *.R script
* Create and load a  package on Github

```{r, echo=FALSE}
rm(list=ls())
EUCOVID <- read.csv("../DATA/COVID-19-ECDCP-2020-08-10.csv")
EUCOVID$dateRep <- gsub("/","-",EUCOVID$dateRep)
EUCOVID$dateRep <- as.Date(EUCOVID$dateRep,format="%m-%d-%Y")
ES <- subset(EUCOVID,geoId =="ES" & cases >0)
ES <- ES[c(1,5:8)]
```

### Plot Spain's Daily Cases And Deaths

```{r}
plot(ES$dateRep,ES$cases,type="b",main="Spain's Daily Cases")
plot(ES$dateRep,ES$deaths,type="b", main="Spain's Daily Deaths")
```
### Part 1 : Simple Non-moving Average
#### create variables of the week and month of each observation:
* https://www.r-bloggers.com/plot-weekly-or-monthly-totals-in-r/

```{r}
ES$Month <- as.Date(cut(ES$dateRep,
  breaks = "month"))
ES$Week <- as.Date(cut(ES$dateRep,
  breaks = "week",
  start.on.monday = FALSE)) # changes weekly break point to Sunday
```
```{r}
ESMC <- aggregate(cases~Week,data=ES,mean)
ESMD <- aggregate(deaths~Week,data=ES,mean)
```
```{r}
plot(ES$dateRep,ES$cases,type="b",main="Spain's Daily and Weekly Mean Cases",col="red")
lines(ESMC$Week,ESMC$cases,col="blue")
```

```{r}
plot(ES$dateRep,ES$deaths,type="b",main="Spain's Daily and Weekly Mean Deaths",col="red")
lines(ESMD$Week,ESMD$deaths,col="blue")
```

### Calculate Monthly Mean Cases and Deaths

```{r}
ESMMC <- aggregate(cases~Month,data=ES,mean)
ESMMD <- aggregate(deaths~Month,data=ES,mean)
```
```{r}
plot(ES$dateRep,ES$cases,type="b",main="Spain's Daily and Monthly Mean Cases",col="red")
lines(ESMMC$Month,ESMMC$cases,col="blue")
```

```{r}
plot(ES$dateRep,ES$deaths,type="b",main="Spain's Daily and Monthly Mean Deaths",col="red")
lines(ESMMD$Month,ESMMD$deaths,col="blue")
```

###  Part 2: Moving Avearge or Rolling Means
In the interest of Full Disclosure, there are a couple Packages that have functions that calculate moving average or rolling mean including: zoo, forecast,pracma and TTR.

